---
title: "IST565_HW4"
author: "Brandon Croarkin"
date: "August 7, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HW 4: Use Clustering to Solve a Mystery in History

In this homework assignment, you are going to use clustering methods to solve a mystery in history: who wrote the disputed essays, Hamilton or Madison?

###Read in Data and Basic EDA

```{r}
library(cluster)
library(wordcloud)
library(ggplot2)
library(ggfortify)
library(dplyr)
```

```{r}
fedPapers <- read.csv('fedPapers85.csv')
head(fedPapers)
```

```{r}
str(fedPapers)
```

```{r}
summary(fedPapers)
```

```{r}
table(fedPapers$author)
```

###Data Preparation

```{r}
#creating a function to find the na's for each column of a dataset
findNAs <- function(df) {
  for(i in names(df)) {
    missingValues = sum(is.na(df[i]))
    print(paste(i, "has" ,missingValues, "missing values."))
  }
}
```

```{r}
findNAs(fedPapers)
```

```{r}
#create a dataframe with just the attribute values
dataNoLabels <- fedPapers[,c(3:ncol(fedPapers))]
head(dataNoLabels)
```

```{r}
#create a dataframe with just the papers from Hamilton, Madison, and the disputed essays
fedPapersHMD <- fedPapers %>% filter(author == "Hamilton" | author == "Madison" | author == "dispt")
table(fedPapersHMD$author)
```

```{r}
#create a dataframe with just the attributes from the Hamilton, Madison, and disputed essays
dataNoLabelsHMD <- fedPapersHMD[,c(3:ncol(fedPapers))]
head(dataNoLabels)
```

###Processing 

```{r}
#make a function to find the most frequent response by group
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```

```{r}
autoplot(prcomp(dataNoLabels), data = fedPapers, colour = 'author', size = 3, alpha = .6)
```

Plot a word cloud for the top ten words from each author. 
```{r}
#create vector of top 20 words
disptWords <- fedPapers %>% filter(author == 'dispt') %>% select(-c("author", "filename")) %>% summarise_all(funs(mean))
disptWords <- sort(disptWords, decreasing = TRUE)

#convert to dataframe
disptWords <- as.matrix(disptWords)
d <- t(disptWords)
colnames(d) <- c("freq")
d <- data.frame(d)
d <- add_rownames(d, "words")


#create wordcloud
wordcloud(words = d$words, freq = d$freq, random.order = FALSE, colors = brewer.pal(8, "Dark2"))
```

```{r}
#create vector of top 20 words
disptWords <- fedPapers %>% filter(author == 'Hamilton') %>% select(-c("author", "filename")) %>% summarise_all(funs(mean))
disptWords <- sort(disptWords, decreasing = TRUE)

#convert to dataframe
disptWords <- as.matrix(disptWords)
d <- t(disptWords)
colnames(d) <- c("freq")
d <- data.frame(d)
d <- add_rownames(d, "words")


#create wordcloud
wordcloud(words = d$words, freq = d$freq, random.order = FALSE, colors = brewer.pal(8, "Dark2"))
```

```{r}
#create vector of top 20 words
disptWords <- fedPapers %>% filter(author == 'Madison') %>% select(-c("author", "filename")) %>% summarise_all(funs(mean))
disptWords <- sort(disptWords, decreasing = TRUE)

#convert to dataframe
disptWords <- as.matrix(disptWords)
d <- t(disptWords)
colnames(d) <- c("freq")
d <- data.frame(d)
d <- add_rownames(d, "words")


#create wordcloud
wordcloud(words = d$words, freq = d$freq, random.order = FALSE, colors = brewer.pal(8, "Dark2"))
```

```{r}
#create vector of top 20 words
disptWords <- fedPapers %>% filter(author == 'Jay') %>% select(-c("author", "filename")) %>% summarise_all(funs(mean))
disptWords <- sort(disptWords, decreasing = TRUE)

#convert to dataframe
disptWords <- as.matrix(disptWords)
d <- t(disptWords)
colnames(d) <- c("freq")
d <- data.frame(d)
d <- add_rownames(d, "words")


#create wordcloud
wordcloud(words = d$words, freq = d$freq, random.order = FALSE, colors = brewer.pal(8, "Dark2"))
```

```{r}
#create vector of top 20 words
disptWords <- fedPapers %>% filter(author == 'HM') %>% select(-c("author", "filename")) %>% summarise_all(funs(mean))
disptWords <- sort(disptWords, decreasing = TRUE)

#convert to dataframe
disptWords <- as.matrix(disptWords)
d <- t(disptWords)
colnames(d) <- c("freq")
d <- data.frame(d)
d <- add_rownames(d, "words")


#create wordcloud
wordcloud(words = d$words, freq = d$freq, random.order = FALSE, colors = brewer.pal(8, "Dark2"))
```

####k-Means (full data) - Trial One

```{r}
set.seed(11)

#choose a k equal to the number of distinct writers
k = 4

#run kmeanModel
(kmeansModel1 <- kmeans(dataNoLabels, k))
```

```{r}
#find cluster sizes
kmeansModel1$size
```

```{r}
#print the centroids
kmeansModel1$centers
```

```{r}
##Get the clusters that were assigned to each group
clusterGroups1 <- data.frame(fedPapers,kmeansModel1$cluster)

#View cluster assignments
View(clusterGroups1[,c(1,ncol(clusterGroups1))])
```

```{r}
#find what author each cluster appears to represent by finding the most common cluster label
df <- clusterGroups1 %>% select(author, kmeansModel1.cluster) %>% group_by(author) %>% mutate(clusterAuthor = Mode(kmeansModel1.cluster))

mytable <-table(df$author,df$kmeansModel1.cluster)
round(prop.table(mytable,1),2)
```

```{r}
set.seed(11)
autoplot(kmeans(dataNoLabels, 4), data = clusterGroups1, colour = 'kmeansModel1.cluster', shape = 'author')
```

```{r}
#use PCA in visualization package "cluster" to visuzalize kMeans model
clusplot(dataNoLabels, kmeansModel1$cluster, color = T, shade = T, labels = 4, lines = 0)
```

```{r}
##plot the clusters
ggplot(data = clusterGroups1, aes(author)) + geom_bar(aes(fill = as.factor(kmeansModel1.cluster)))
```

####k-Means (full data) - Trial Two

```{r}
set.seed(1)

#choose a k equal to the number of distinct writers
k = 4

#run kmeanModel
(kmeansModel2 <- kmeans(dataNoLabels, k))
```

```{r}
#find cluster sizes
kmeansModel2$size
```

```{r}
#print the centroids
kmeansModel2$centers
```

```{r}
##Get the clusters that were assigned to each group
clusterGroups2 <- data.frame(fedPapers,kmeansModel2$cluster)

#View cluster assignments
View(clusterGroups2[,c(1,ncol(clusterGroups2))])
```

```{r}
set.seed(1)
autoplot(kmeans(dataNoLabels, 4), data = dataNoLabels)
```

```{r}
#find what author each cluster appears to represent by finding the most common cluster label
df <- clusterGroups2 %>% select(author, kmeansModel2.cluster) %>% group_by(author) %>% mutate(clusterAuthor = Mode(kmeansModel2.cluster))

mytable <-table(df$author,df$kmeansModel2.cluster)
round(prop.table(mytable,1),2)
```

```{r}
set.seed(11)
autoplot(kmeans(dataNoLabels, 4), data = clusterGroups2, colour = 'kmeansModel2.cluster', shape = 'author')
```

```{r}
#use PCA in visualization package "cluster" to visuzalize kMeans model
clusplot(dataNoLabels, kmeansModel2$cluster, color = T, shade = T, labels = 4, lines = 0)
```

```{r}
##plot the clusters
ggplot(data = clusterGroups2, aes(author)) + geom_bar(aes(fill = as.factor(kmeansModel2.cluster)))
```

####k-Means (full data) - Trial Three

```{r}
set.seed(65)

#choose a k equal to the number of distinct writers
k = 4

#run kmeanModel
(kmeansModel3 <- kmeans(dataNoLabels, k))
```

```{r}
#find cluster sizes
kmeansModel3$size
```

```{r}
#print the centroids
kmeansModel3$centers
```

```{r}
##Get the clusters that were assigned to each group
clusterGroups3 <- data.frame(fedPapers,kmeansModel3$cluster)

#View cluster assignments
View(clusterGroups3[,c(1,ncol(clusterGroups3))])
```

```{r}
set.seed(65)
autoplot(kmeans(dataNoLabels, 4), data = dataNoLabels)
```

```{r}
#find what author each cluster appears to represent by finding the most common cluster label
df <- clusterGroups3 %>% select(author, kmeansModel3.cluster) %>% group_by(author) %>% mutate(clusterAuthor = Mode(kmeansModel3.cluster))

mytable <-table(df$author,df$kmeansModel3.cluster)
round(prop.table(mytable,1),2)
```

```{r}
set.seed(11)
autoplot(kmeans(dataNoLabels, 4), data = clusterGroups3, colour = 'kmeansModel3.cluster', shape = 'author')
```

```{r}
#use PCA in visualization package "cluster" to visuzalize kMeans model
clusplot(dataNoLabels, kmeansModel3$cluster, color = T, shade = T, labels = 4, lines = 0)
```

```{r}
##plot the clusters
ggplot(data = clusterGroups3, aes(author)) + geom_bar(aes(fill = as.factor(kmeansModel3.cluster)))
```

####k-Means (subset data) - Trial Four

```{r}
set.seed(11)

#choose a k equal to the number of distinct writers
k = 2

#run kmeanModel
(kmeansModel4 <- kmeans(dataNoLabelsHMD, k))
```

```{r}
#find cluster sizes
kmeansModel4$size
```

```{r}
#print the centroids
kmeansModel4$centers
```

```{r}
##Get the clusters that were assigned to each group
clusterGroups4 <- data.frame(fedPapersHMD,kmeansModel4$cluster)

#View cluster assignments
View(clusterGroups4[,c(1,ncol(clusterGroups4))])
```

```{r}
set.seed(11)
autoplot(kmeans(dataNoLabelsHMD, 2), data = dataNoLabelsHMD)
```

```{r}
#find what author each cluster appears to represent by finding the most common cluster label
df <- clusterGroups4 %>% select(author, kmeansModel4.cluster) %>% group_by(author) %>% mutate(clusterAuthor = Mode(kmeansModel4.cluster))

mytable <-table(df$author,df$kmeansModel4.cluster)
round(prop.table(mytable,1),2)
```

```{r}
#use PCA in visualization package "cluster" to visuzalize kMeans model
clusplot(dataNoLabelsHMD, kmeansModel4$cluster, color = T, shade = T, labels = 2, lines = 0)
```

```{r}
##plot the clusters
ggplot(data = clusterGroups4, aes(author)) + geom_bar(aes(fill = as.factor(kmeansModel4.cluster)))
```

####k-Means (subset data) - Trial Five

```{r}
set.seed(1)

#choose a k equal to the number of distinct writers
k = 2

#run kmeanModel
(kmeansModel5 <- kmeans(dataNoLabelsHMD, k))
```

```{r}
#find cluster sizes
kmeansModel5$size
```

```{r}
#print the centroids
kmeansModel5$centers
```

```{r}
##Get the clusters that were assigned to each group
clusterGroups5 <- data.frame(fedPapersHMD,kmeansModel5$cluster)

#View cluster assignments
View(clusterGroups5[,c(1,ncol(clusterGroups5))])
```

```{r}
set.seed(1)
autoplot(kmeans(dataNoLabelsHMD, 2), data = dataNoLabelsHMD)
```

```{r}
#use PCA in visualization package "cluster" to visuzalize kMeans model
clusplot(dataNoLabelsHMD, kmeansModel5$cluster, color = T, shade = T, labels = 2, lines = 0)
```

```{r}
#find what author each cluster appears to represent by finding the most common cluster label
df <- clusterGroups5 %>% select(author, kmeansModel5.cluster) %>% group_by(author) %>% mutate(clusterAuthor = Mode(kmeansModel5.cluster))

mytable <-table(df$author,df$kmeansModel5.cluster)
round(prop.table(mytable,1),2)
```

```{r}
##plot the clusters
ggplot(data = clusterGroups5, aes(author)) + geom_bar(aes(fill = as.factor(kmeansModel5.cluster)))
```

####k-Means (subset data) - Trial Six

```{r}
set.seed(65)

#choose a k equal to the number of distinct writers
k = 2

#run kmeanModel
(kmeansModel6 <- kmeans(dataNoLabelsHMD, k))
```

```{r}
#find cluster sizes
kmeansModel6$size
```

```{r}
#print the centroids
kmeansModel6$centers
```

```{r}
##Get the clusters that were assigned to each group
clusterGroups6 <- data.frame(fedPapersHMD,kmeansModel6$cluster)

#View cluster assignments
View(clusterGroups6[,c(1,ncol(clusterGroups6))])
```

```{r}
set.seed(65)
autoplot(kmeans(dataNoLabelsHMD, 2), data = dataNoLabelsHMD)
```

```{r}
#use PCA in visualization package "cluster" to visuzalize kMeans model
clusplot(dataNoLabelsHMD, kmeansModel6$cluster, color = T, shade = T, labels = 2, lines = 0)
```

```{r}
#find what author each cluster appears to represent by finding the most common cluster label
df <- clusterGroups6 %>% select(author, kmeansModel6.cluster) %>% group_by(author) %>% mutate(clusterAuthor = Mode(kmeansModel6.cluster))

mytable <-table(df$author,df$kmeansModel6.cluster)
round(prop.table(mytable,1),2)
```

```{r}
##plot the clusters
ggplot(data = clusterGroups6, aes(author)) + geom_bar(aes(fill = as.factor(kmeansModel6.cluster)))
```

```{r}
centers <- clusterGroups6 %>% select(-c(author, filename)) %>% group_by(kmeansModel6.cluster) %>% summarize_all(funs(mean))
centers <- t(centers)
colnames(centers) <- c("Cluster1", "Cluster2")
centers <- centers[-1,]
centers <- round(centers, 3)
centers <- data.frame(centers)
centers$diff <- abs(centers$Cluster1 - centers$Cluster2)
centers <- add_rownames(centers, "FunctionWord")
centers <- centers %>% arrange(desc(diff)) %>% top_n(20)
centers
```

```{r}

```

```{r}

```

#####HAC

#####Full Dataset

######Creating Distance Measures

First step involves making different distance measures. 
```{r}
distMatrix_E <- dist(dataNoLabels, method="euclidean") 
```

```{r}
distMatrix_M <- dist(dataNoLabels, method="manhattan")
```

```{r}
distMatrix_C <- dist(dataNoLabels, method="canberra")
```

Then we create the hierarchical clusters.

######Euclidean Distance

#######Complete Linkage
```{r}
hc_e <- hclust(distMatrix_E, method = "complete")
plot(hc_e)
```

```{r}
clust_e <- cutree(hc_e, k=4)
with(fedPapers, tapply(clust_e, list(author), mean))
```

```{r}
table(clust_e, fedPapers$author)
```

#######Average Linkage

```{r}
hc_e2 <- hclust(distMatrix_E, method = "average")
plot(hc_e2)
```

```{r}
clust_e2 <- cutree(hc_e2, k=4)
with(fedPapers, tapply(clust_e2, list(author), mean))
```

```{r}
table(clust_e2, fedPapers$author)
```

#######Single Linkage

```{r}
hc_e3 <- hclust(distMatrix_E, method = "single")
plot(hc_e3)
```

```{r}
clust_e3 <- cutree(hc_e3, k=4)
with(fedPapers, tapply(clust_e3, list(author), mean))
```

```{r}
table(clust_e3, fedPapers$author)
```

######Manhattan Distance

#######Complete Linkage
```{r}
hc_m <- hclust(distMatrix_M, method = "complete")
plot(hc_m)
```

```{r}
clust_e <- cutree(hc_m, k=4)
with(fedPapers, tapply(clust_e, list(author), mean))
```

```{r}
table(clust_e, fedPapers$author)
```

#######Average Linkage

```{r}
hc_m2 <- hclust(distMatrix_M, method = "average")
plot(hc_m2)
```

```{r}
clust_e2 <- cutree(hc_m2, k=4)
with(fedPapers, tapply(clust_e2, list(author), mean))
```

```{r}
table(clust_e2, fedPapers$author)
```

#######Single Linkage

```{r}
hc_m3 <- hclust(distMatrix_M, method = "single")
plot(hc_m3)
```

```{r}
clust_e3 <- cutree(hc_m3, k=4)
with(fedPapers, tapply(clust_e3, list(author), mean))
```

```{r}
table(clust_e3, fedPapers$author)
```

######Canberra Distance

#######Complete Linkage
```{r}
hc_c <- hclust(distMatrix_C, method = "complete")
plot(hc_c)
```

```{r}
clust_e <- cutree(hc_c, k=4)
with(fedPapers, tapply(clust_e, list(author), mean))
```

```{r}
table(clust_e, fedPapers$author)
```

#######Average Linkage

```{r}
hc_c2 <- hclust(distMatrix_C, method = "average")
plot(hc_c2)
```

```{r}
clust_e2 <- cutree(hc_c2, k=4)
with(fedPapers, tapply(clust_e2, list(author), mean))
```

```{r}
table(clust_e2, fedPapers$author)
```

#######Single Linkage

```{r}
hc_c3 <- hclust(distMatrix_C, method = "single")
plot(hc_c3)
```

```{r}
clust_e3 <- cutree(hc_c3, k=4)
with(fedPapers, tapply(clust_e3, list(author), mean))
```

```{r}
table(clust_e3, fedPapers$author)
```

#####Hamilton and Madison Subset

######Creating Distance Measures

First step involves making different distance measures. 
```{r}
distMatrix_E2 <- dist(dataNoLabelsHMD, method="euclidean")
```

```{r}
distMatrix_M2 <- dist(dataNoLabelsHMD, method="manhattan")
```

```{r}
distMatrix_C2 <- dist(dataNoLabelsHMD, method="canberra")
```

Then we create the hierarchical clusters.

######Euclidean Distance

#######Complete Linkage
```{r}
hc_e <- hclust(distMatrix_E2, method = "complete")
plot(hc_e)
```

```{r}
clust_e <- cutree(hc_e, k=2)
with(fedPapersHMD, tapply(clust_e, list(author), mean))
```

```{r}
table(clust_e, fedPapersHMD$author)
```

#######Average Linkage

```{r}
hc_e2 <- hclust(distMatrix_E2, method = "average")
plot(hc_e2)
```

```{r}
clust_e2 <- cutree(hc_e2, k=2)
with(fedPapersHMD, tapply(clust_e2, list(author), mean))
```

```{r}
table(clust_e2, fedPapersHMD$author)
```

#######Single Linkage

```{r}
hc_e3 <- hclust(distMatrix_E2, method = "single")
plot(hc_e3)
```

```{r}
clust_e3 <- cutree(hc_e3, k=2)
with(fedPapersHMD, tapply(clust_e3, list(author), mean))
```

```{r}
table(clust_e3, fedPapersHMD$author)
```

######Manhattan Distance

#######Complete Linkage
```{r}
hc_m <- hclust(distMatrix_M2, method = "complete")
plot(hc_m)
```

```{r}
clust_e <- cutree(hc_m, k=2)
with(fedPapersHMD, tapply(clust_e, list(author), mean))
```

```{r}
table(clust_e, fedPapersHMD$author)
```

#######Average Linkage

```{r}
hc_m2 <- hclust(distMatrix_M2, method = "average")
plot(hc_m2)
```

```{r}
clust_e2 <- cutree(hc_m2, k=2)
with(fedPapersHMD, tapply(clust_e2, list(author), mean))
```

```{r}
table(clust_e2, fedPapersHMD$author)
```

#######Single Linkage

```{r}
hc_m3 <- hclust(distMatrix_M2, method = "single")
plot(hc_m3)
```

```{r}
clust_e3 <- cutree(hc_m3, k=2)
with(fedPapersHMD, tapply(clust_e3, list(author), mean))
```

```{r}
table(clust_e3, fedPapersHMD$author)
```

######Canberra Distance

#######Complete Linkage
```{r}
hc_c <- hclust(distMatrix_C2, method = "complete")
plot(hc_c)
```

```{r}
clust_e <- cutree(hc_c, k=2)
with(fedPapersHMD, tapply(clust_e, list(author), mean))
```

```{r}
table(clust_e, fedPapersHMD$author)
```

#######Average Linkage

```{r}
hc_c2 <- hclust(distMatrix_C2, method = "average")
plot(hc_c2)
```

```{r}
clust_e2 <- cutree(hc_c2, k=2)
with(fedPapersHMD, tapply(clust_e2, list(author), mean))
```

```{r}
table(clust_e2, fedPapersHMD$author)
```

#######Single Linkage

```{r}
hc_c3 <- hclust(distMatrix_C2, method = "single")
plot(hc_c3)
```

```{r}
clust_e3 <- cutree(hc_c3, k=2)
with(fedPapersHMD, tapply(clust_e3, list(author), mean))
```

```{r}
table(clust_e3, fedPapersHMD$author)
```




